Thingies for Dummies:\\A smart home infrastructure for the rest of us

### 2012-05-17 report/paper.tex
pdf figures
report finished

\documentclass{ubicomp2012}
\usepackage{times}
\usepackage{url}
\usepackage{graphics}
\usepackage{color}
\usepackage{mathtools}
\usepackage{float}
\usepackage[pdftex]{hyperref}
\usepackage{listings}
\lstset{ %
  % language=C,             % the language of the code
  basicstyle=\ttfamily,         % the size of the fonts that are used for the code
  showspaces=false,             % show spaces adding particular underscores
  showstringspaces=false,       % underline spaces within strings
  showtabs=false,              % show tabs within strings adding particular underscores
  tabsize=2,                    % sets default tabsize to 2 spaces
  captionpos=b,                 % sets the caption-position to bottom
  breaklines=true,              % sets automatic line breaking
  breakatwhitespace=false     % sets if automatic breaks should only happen at witespace
}
\hypersetup{%
pdftitle={Thinges for Dummies - A Smart}, pdfauthor={David Thomas, Egil Hansen}, pdfkeywords={your
keywords}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
breaklinks=true, }
\newcommand{\comment}[1]{}
\definecolor{Orange}{rgb}{1,0.5,0}
\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[#1]]}}}}

@@

\title{Thingies for Dummies\\A smart home infrastructure for the rest of us}
\numberofauthors{2}
\author{
  \alignauthor David Thomas\\
    \email{dtho@itu.dk}
 \alignauthor Egil Hansen\\
    \email{ekri@itu.dk}  }
\maketitle

\begin{abstract}

\end{abstract}

\keywords{TODO.}

\category{H.5.2}{Information interfaces and presentation (e.g., HCI)}{Miscellaneous}.

\generalterms{Design,
Documentation, Economics, Experimentation, Human Factors, Languages, Management, Measurement, Performance, Reliability,
Security, Standardization, Theory, Verification.}

@@

More than 20 years after Weiser first defined ubiquitous computing \cite{Weiser1991} we have yet to see smart homes become commonplace. We believe this is largely due to smart home technology being inaccessible and too complicated for everyday users, and both price, perceived lack of advantages, and challenges in learning new technology scare users away.

In this paper we propose a smart home infrastructure that allow homeowners to cheaply bring new smart home technology into their existing homes in small batches, technology that homeowners will realistically be able to install and manage themselves. Our proposed infrastructure bases itself on hardware that exists in most homes today -- Wi-Fi and smart phones -- and uses simple and open protocols like HTTP and the WebSocket protocol to provide the communication backbone of the infrastructure. Each individual device, or ``Thingy'', as we have dubbed them, is supposed to be simple to install and use, but provide value nonetheless. We call this concept \textit{Thingies for Dummies}\footnote{This is a reference to the \textit{For Dummies} book series that tries to present non-intimidating guides to various topics and that always contains the subtitle ``A Reference for the Rest of Us!''.}.

Edwards and Grinter describe, in their 2001 paper \cite{Edwards01athome}, seven challenges facing the adoption of smart home technology\footnote{1: \textit{The `Accidentally' Smart Home}. 2: \textit{Impromptu Interoperability (islands of functionality)}. 3: \textit{No Systems Administrator}. 4: \textit{Designing for Domestic Use}. 5: \textit{Social Implications of Aware Home Technologies}. 6: \textit{Reliability}. 7: \textit{Inference in the Presence of Ambiguity}.}, and we believe that through the overarching theme of `simple' in our proposal -- simple infrastructure and simple thingies -- we are able to answer the first six challenges, at least partly. Simplifying does imply giving up on the dream of a fully automated intelligent home for the time being, but our proposed infrastructure does not prevent a home from becoming gradually more automated or intelligent as developers become better at leveraging the platform.

When we started doing research for this project we first wanted to see if we could come up with a few use cases where some smart home technology would add so much value, that a normal homeowner would go out and invest time and money to retrofit the home. Through our informal questioning of friends, colleagues, and family, it quickly became clear that the promise of the classic intelligent appliances we see in many office buildings that automatically adjusts their operations through context awareness\footnote{Appliances such as HVAC, light fixtures, window curtains, and automatic doors.} did not have a strong enough appeal. The only thing that really resonated was various remote control scenarios. Examples included double checking if all the lights are turned off at home, controlling radiators, remotely starting the washer when leaving from work so clean clothes are ready to be hung up to dry when the user gets home, and remotely unlocking the front door to the home if a friend is waiting and the owner is running late. With that in mind,  we created the following scenario to guide our design process:

@@

\section{Thinges for Dummies Infrastructure}
The key design decision for \textit{thingies for Dummies} is that it should be simple and  inexpensive to get started retrofitting an existing house with thingies and that it should be independent of specific hardware and software platforms. This means not requiring a dedicated, centralized controller, and instead to rely on the existing infrastructure in the home for handling communication and control, i.e. a Wi-Fi network and smart phones. By doing so, we limit the amount of new hardware required to get started, since most homes will already have a Wi-Fi network and at least one smartphone. This lowers the price and reduces the environmental impact. Figure \ref{fig:infrastructureOverview} shows an overview of the infrastructure:

\begin{figure}[th]
\begin{center}
\includegraphics[width=0.90\columnwidth]{figures/systemoverviewdrawing.png}
\end{center}
\caption{Overview of the infrastructure}
\label{fig:infrastructureOverview}
\end{figure}

An existing Wi-Fi network in the home provides the communications channel. The user's smartphone is connected to this network. A number of thingies -- exemplified in the figure as a light switch and a window sensor -- are brought into the home and connected to the Wi-Fi network as well. An app on the smartphone then enables the user control and check on the thingies.

@@

With so many networking technologies available care must be taken before picking one above another, however, we arrived at our choice, Wi-Fi, pretty fast. One of our primary requirements was that the network should exist in most homes already. This eliminates quite a few choices right away, leaving us with power lines, which the X10 protocol uses, cellular networks such as 3G, phone line wiring, Bluetooth, and Wi-Fi. Power lines, which has the advantage of reaches all rooms in a home, has problems with the an uncertain network boundaries, since there is usually no special gateway or firewall preventing data from traveling on the power lines to the neighbours power lines, unless extra hardware is added. Power lines are also inherently a wired network, which means an extra bridge device is needed between the power lines and wireless mobile devices if those are to interact with devices connected to it. The phone lines inside the home has similar problems, and is also far less ubiquitous in the home, with most homes only having a few phone plugs. Then there is Bluetooth, which most laptops and phones, even feature phones, come equipped with. Bluetooth is however limited in range. This leaves cellular networks, which are indeed wireless, but also owned and maintained by carrier companies, that charges fees for the data that travels on their network and not all areas are covered equally well. This makes cellular networks the less ideal choice, landing us squarely on Wi-Fi as the preferred option.

@@

The biggest drawback related to Wi-Fi is that it is power hungry compared to other technologies such as ZigBee \cite{zigbeeWebsite}. However, there is a general focus from the Wi-Fi radio chip makers to lower power consumption due to the massive surge of battery driven mobile devices the last few years, so the drawback will matter less as Wi-Fi radios become more power friendly. In addition, for many of the scenarios we have found users to be interested in, thingies will have a permanent power source and in the case where low powered sensor thingies are required, we expect these to use a protocol like ZigBee\cite{zigbeeWebsite} and be bundled with a special purpose thingy which act as a bridge between the individual ZigBee enabled thingies the Wi-Fi network. The bridge thingy could also perform other tasks so the extra cost would be negligible.

We also discussed the problems with oversaturation and general instabilities of the average home Wi-Fi network. This is of course very dependent on the model of the wireless access point used. Better models with good Wi-Fi radios, antennas, faster CPUs, and enough memory, can handle around 25 concurrent connected clients, while lesser models choke at at half as many connections. However, since one of our design goals was that it should be as inexpensive as possible to get started with smart home technology, we do not expect a average home Wi-Fi to get saturated right away. The scenario where a homeowner invests in over 20 new thingies at the same time might not be unlikely, but in that case we feel that investment in an extra access point to handle traffic from thingies is probably also within the budget. It is common knowledge among Some Wi-Fi experts claim that Wi-Fi clients also have a tendency to drop their connection from access points more often if using the higher bandwidth 802.11n and 802.11g protocols compared to the slower 802.11b. The solution here is to make thingies use 802.11b where possible, and since only a very small subset of thingies will require a 100\% stable connection, homeowners should experience no problems since thingies can simply reconnect if their connection drops.
\subsubsection{Communication protocols}

For communication, we use standard HTTP \cite{Fielding:1999:HTP:RFC2616}. This opens up for a huge list of possible controllers, basically any type of computing device or library that supports HTTP. This is especially important to us since we do not want to be limited to a certain framework, language, or platform, e.g. Java\footnote{\url{http://java.com}} or Android\footnote{\url{http://android.com}}. Other great properties of HTTP is that it is fairly lightweight and stateless protocol \cite{Fielding:1999:HTP:RFC2616}, which simplifies things considerably on the thingies. HTTP also supports our goal of an open infrastructure; it might very well be the best supported protocol among developers, and this enables the average app developer to build new functionality on top of the infrastructure on their platform of choice. Another benefit is that HTTP communication is very unlikely to be blocked in firewalls compared to other less used protocols, so there is a much bigger chance that things will just work.

However, HTTP's statelessness is also a weakness in common smart home scenarios, where a context change in one thingy should trigger an event in another thingy. With basic HTTP, a thingy (client) waiting for a state change has to continuously poll the other thingy it is monitoring to discover the state change in a timely manner. This is an waste of resources and could potentially deplete the resources of a thingy entirely, especially if more clients are waiting for changes. Clearly, relying only on HTTP alone does not scale. Several solutions to this problem were considered, the biggest contender to be considered being multicast. The advantages of using multicast is that thingies can simply announce state changes on the network and it would not matter if zero, one, or hundreds of clients were interested, the amount of resources required to do so would be the same. Ignoring the problems associated multicast on Wi-Fi, i.e. lost multicast packets are not retransmitted at the MAC layer \cite{conf/icnp/ChandraKMNPRR09}, the reason we did not choose multicast was that the protocol works on a lower level compared to HTTP, and that would limit the number of supported development platforms. We wanted to stay in the Web-stack of technologies so we chose the WebSocket protocol. In the recent years, the WebSocket protocol has been standardized by the Internet Engineering Task Force (IETF) \cite{IETFWebsockets} and is in the process of being standardized by the World Wide Web Consortium (W3C) \cite{w3cWebsockets}. The WebSocket protocol supports multiplexing bi-directional, full-duplex communications channels over a single TCP connection and is already supported by the latests versions of most major browsers\footnote{The WebSocket protocol is supported from Chrome 16, Firefox 11 and Internet Explorer 10.}, with numerous frameworks and libraries already built and more on the way. The WebSocket Protocol was essentially designed to be a solution for the scenario where a server needs to push information to a client\cite{IETFWebsockets} which web developers have been struggling with for many years, so while it might not scale as well as multicast, we feel this is the best overall approach.

\subsubsection{Service Discovery}
Most home networks do not have traditional unicast DNS server set up, so we need a different way to discover thingies connect to the network, both during installation and also if and when their IP address changes. We have decided to use mDNS/Bonjour\footnote{Bonjour is available on \url{https://developer.apple.com/opensource/}} since it offers wide support across platforms and is relatively lightweight and open source. .
\subsubsection{Security considerations}

The basic trust model for thingies for Dummies is based on the security of the home's Wi-Fi. Assuming the Wi-Fi is probably configured using WPA2 with AES encryption and a reasonable passphrase, there is a clear boundary for the smart home, and an adversary would not be able to pick up data sent to and from thingies without being connected to the network. This also means, that anyone on the network have full control over the thingies attached to it, which can be a problem in some cases. We have discussed the possibility of adding an extra layer of security that would handle such a case. If the concern is only that unauthorised users cannot perform certain actions on a thingy, i.e. only change state, not update configuration, then we need a way to authenticate each user. This could be done using certificates that clients use to authenticate with. The certificate would be generated during the initial configuration of a thingy, and depending on the number of authorisation levels needed for a thingy, e.g. read only, update state, and update configuration, one or more unique certificates would be generated. Then it would be up to the controller devices used during the initial configuration to distribute the generated certificates to authorized users. However, it is not enough only to authenticate users, since the communication between users and thingies are still being transmitted over an open channel which can allow an adversary to perform replay attacks. The solution to this is to secure the communication channel between users and Thinges, i.e. using HTTPS. Adding the extra level of security does raise the hardware requirements on thingies which increases cost and power consumption, so an environment that supports both ``open'', low risk thingies and ``secure'' thingies should provide the best of both worlds.

\subsection{thingies - actuators and sensors}

@@

The first two points are important because thingies must be installable and manageable by non-technical users. It is also reasonable to assume that a simple thingy with one or two functions is going to be more reliable than a device with several functions. Not allowing thingies to be dependent on other thingies or services simplifies everything, which again helps makes things easier for the non-technical user. Making things simple also means being less dependent on constant firmware updates, which will limit the amount of maintenance required after the initial installation. This is highly important, as Kindberg et al writes, \textit{ ``You can't reboot the world, let alone rewrite it, to introduce new functionality. Users shouldn't be led into a disposable physical world just because they can't keep up with software upgrades.''} \cite{kindbergSystem}.

thingies can have running logic inside them, depending on their functionality, e.g. a power plug could be configured to turn on and off at specific times during the day, or turn on a secondary output if a specified amount of power is being drawn from a primary output, but it cannot depend on context outside of its domain, i.e. other thingies. Such logic should exist on controllers.

@@

\begin{itemize}
\item Have a unique identifier
\item Have a passive NFC tag and/or QR code with its ID
\item Have a physical reset/install mode button
\item Be able to connect to an IP based network
\end{itemize}

@@

Why not require a Wi-Fi radio in thingies? This is covered by the last requirement, but is more generalized, since we want to allow for thingies that connect directly through the LAN interface on the home Wi-Fi router, and also enable the scenario where a ZigBee enabled sensor thingies connect through a specialized ZigBee to Wi-Fi bridge thingy. We do however still require that all thingies can be reached on their own unique URI through the network. This is trivial for thingies connecting directly through Wi-Fi or LAN to the home network -- they simply have a standard URI \cite{uri-rfc} based on their IP address, e.g. \lstinline{http://192.168.0.42}. In the bridge scenario, the bridged thingies, e.g. ZigBee sensors, may have to share the IP of the bridge. Luckily, the URI schema allows for ports to be defined, so each bridged thingy will be identified by the bridges IP and their own port, e.g. \lstinline{http://192.168.0.42:1337}, \lstinline{http://192.168.0.42:1338}, etc.

@@

The structured description associates each state variable in the thingy with an internally unique ID. This is what makes it possible to keep the service interface simple, as these IDs are used when querying and setting state. The following list summarizes the web service interface:

@@

The general bootstrapping process from a user's point of view is very simple. First, the user plugs in and turns on the thingy. Then, the user starts the Home Remote app on his smartphone and, after entering credentials for the Wi-Fi network, clicks ``Install new Thingy''. The onscreen instructions will prompt the user to scan the QR code or hold the phone close to the thingy (if the phone is NFC enabled). From this point, the phone and the thingy are connected to each other, and the user is asked to give it a name so it is identifiable by humans. The phone also sends the Wi-Fi credentials to the thingy. The diagram in figure \ref{fig:installworkflow} illustrates the workflow of the bootstrapping protocol from a technical point of view.

@@

You cannot have a smart home without controllers, and to us, the obvious first choice for controllers are smart phones. It is a device which users carry all the time and are comfortable using. As the previous sections show, we are by no means limited to smart phones -- any computing device supporting the HTTP protocol can act as a controller. This is also our only requirements for simple controllers, although we do require a full controller to be able to perform bootstrapping of new thingies, which is where smartphones really shine.
To enable the simplest form of user programming and automation on the controller level, and to make the scenario described in the introduction possible, we need a way to execute a batch of commands to different thingies. Our solution to this is something we have dubbed ``Presets''. A Preset consists of a name and a set of thingies, along with a desired state for each thingy. When the user defines a new Preset, he gives it a name and then adds one thingy at a time to the preset, specifying which state it should be in.When a Preset is triggered by the user, all thingies in the preset are put in the required state in sequential order, and the expected state of each thingies is checked. If any thingies in the Preset did not end up in the expected state, the user is notified. This also applies to thingies that are only sensors, where the Preset is simply used to verify their states. This enables relatively advanced scenarios such as our ``Lock Down'' scenario described earlier.

@@

The Internet of Things, implemented in e.g. HP's Cooltown project\cite{cooltown} is a concept related to ours in that it connects the physical and virtual worlds by creating an online presence for physical objects. The Cooltown project also mentions the use of local wireless networks as a means of access restriction.

\subsection{Smart home technologies}

X10\footnote{\url{http://x10.com}} is a low-level protocol from the 70s running through the power lines in a home. It allows one to control a number of devices using a pre-defined set of commands \footnote{url{http://software.x10.com/pub/manuals/xtdcode.pdf}}. Thus, it requires a controller device which is connected to the power lines. The power line medium also introduces potential issues with X10 commands leaving or entering a home.

The Open Services Gateway Initiative, OSGi\cite{osgiIntroduction}, is an infrastructure designed by an alliance of over 80 companies. \cite{osgiDiscovery} describes how the technology can be used in the context of a smart home, including integration with devices through other technologies such as UPnP and Jini. The approach makes use of a gateway device within the home network. The infrastructure is marketed towards vendors and used in various products and solutions\cite{osgiMarkets}.

>>>>

\documentclass{ubicomp2012}
\usepackage{times}
\usepackage{url}
\mdseries\itshape\urlstyle{same}
\usepackage{graphics}
\usepackage{color}
\usepackage{mathtools}
\usepackage{float}
\usepackage[pdftex]{hyperref}
\usepackage{listings}
\lstset{ %
  % language=C,             % the language of the code
  basicstyle=\ttfamily,         % the size of the fonts that are used for the code
  showspaces=false,             % show spaces adding particular underscores
  showstringspaces=false,       % underline spaces within strings
  showtabs=false,              % show tabs within strings adding particular underscores
  tabsize=2,                    % sets default tabsize to 2 spaces
  captionpos=b,                 % sets the caption-position to bottom
  breaklines=true,              % sets automatic line breaking
  breakatwhitespace=false     % sets if automatic breaks should only happen at witespace
}
\hypersetup{%
pdftitle={Thinges for Dummies},
pdfauthor={David Thomas, Egil Hansen},
pdfkeywords={UI-less devices, smart homes, remote control, infrastructure.},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
breaklinks=true, }
\newcommand{\comment}[1]{}
\definecolor{Orange}{rgb}{1,0.5,0}
\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[#1]]}}}}

@@

\title{Thingies for Dummies:\\A smart home infrastructure for the rest of us}
\numberofauthors{2}
\author{
  \alignauthor David Thomas\\
    \affaddr{IT University of Copenhagen}\\
    \email{dtho@itu.dk}
 \alignauthor Egil Hansen\\
    \affaddr{IT University of Copenhagen}\\
    \email{ekri@itu.dk}  }
\maketitle

\begin{abstract}
A number of smart home infrastructures and technologies exist. However, these are not commonly adopted by homeowners as they are often too expensive, complex and difficult to retrofit in existing homes. We introduce an infrastructure that enables end-users to bring simple smart home devices into their homes and install them. Our infrastructure makes use of equipment already present in many homes -- Wi-Fi networks and smartphones -- and is based on common web technologies. We include a bootstrapping process to connect UI-less devices to Wi-Fi networks, and an approach to generate user interfaces for these devices.
We have evaluated our designed infrastructure via a user test of an implemented prototype, and our evaluation participants found the prototype easy to install and use overall.
\end{abstract}

\keywords{UI-less devices, smart homes, remote control, infrastructure.}

%\category{H.5.2}{Information interfaces and presentation (e.g., HCI)}{Miscellaneous}.

\generalterms{Design, Human Factors, Security.}

@@

More than 20 years after Weiser first defined ubiquitous computing \cite{Weiser1991}, we have yet to see smart homes become commonplace. We believe this is largely due to smart home technology being inaccessible and too complicated for everyday users, and both price, perceived lack of advantages, and challenges in learning new technology scare users away.

In this paper we propose a smart home infrastructure that allow homeowners to cheaply bring new smart home technology into their existing homes in small batches, technology that homeowners will realistically be able to install and manage themselves. Our proposed infrastructure bases itself on hardware that exists in many homes today -- Wi-Fi and smartphones -- and uses simple and open protocols like HTTP and the WebSocket protocol to provide the communication backbone of the infrastructure. Each individual device, or ``thingy'', as we have dubbed them, is supposed to be simple to install and use, but provide value nonetheless. We call this concept \textit{Thingies for Dummies}\footnote{This is a reference to the \textit{For Dummies} book series that tries to present non-intimidating guides to various topics and that always contains the subtitle ``A Reference for the Rest of Us!''.}.

Edwards and Grinter describe, in their 2001 paper \cite{Edwards01athome}, seven challenges facing the adoption of smart home technology:
\begin{enumerate}
\item The `Accidentally' Smart Home
\item Impromptu Interoperability (islands of functionality)
\item No Systems Administrator
\item Designing for Domestic Use
\item Social Implications of Aware Home Technologies
\item Reliability
\item Inference in the Presence of Ambiguity
\end{enumerate}

We believe that through the overarching theme of `simple' in our proposal -- simple infrastructure and simple thingies -- we are able to answer the first six challenges, at least in part. Simplifying does imply giving up on the idea of a fully automated intelligent home for the time being, but our proposed infrastructure does not prevent a home from becoming gradually more automated or intelligent as developers become better at leveraging the platform.

\subsection{Do people want smart homes?}
When we started doing research for this project, we first wanted to see if we could come up with a few use cases where some smart home technology would add so much value, that a normal homeowner would go out and invest time and money to retrofit the home. Through our informal questioning of friends, colleagues, and family, it quickly became clear that the promise of the classic intelligent appliances we see in many office buildings that automatically adjusts their operations through context awareness\footnote{Appliances such as HVAC, light fixtures, window curtains, and automatic doors.} did not have a strong enough appeal. The only thing that really resonated was various remote control scenarios. Examples included double checking if all the lights are turned off at home, controlling radiators, remotely starting the washer when leaving from work so clean clothes are ready to be hung up to dry when the user gets home, and remotely unlocking the front door to the home if a friend is waiting and the owner is running late. With that in mind,  we created the following scenario to guide our design process:

@@

\section{Thinges for Dummies Infrastructure}
The key design decision for \textit{Thingies for Dummies} is that it should be simple and  inexpensive to get started retrofitting an existing house with thingies and that it should be independent of specific hardware and software platforms. This means not requiring a dedicated, centralized controller, and instead to rely on the existing infrastructure in the home for handling communication and control, i.e. a Wi-Fi network and smartphones. By doing so, we limit the amount of new hardware required to get started, since most homes will already have a Wi-Fi network and at least one smartphone. This lowers the price and reduces the environmental impact. Figure \ref{fig:infrastructureOverview} shows an overview of the infrastructure.

\begin{figure}[th]
\begin{center}
\includegraphics[width=0.90\columnwidth]{figures/systemoverviewdrawing.pdf}
\end{center}
\caption{Overview of the infrastructure}
\label{fig:infrastructureOverview}
\end{figure}

An existing Wi-Fi network in the home provides the communications channel. The user's smartphone is connected to this network. A number of thingies -- exemplified in the figure as a light switch and a window sensor -- are brought into the home and connected to the Wi-Fi network as well. An app on the smartphone then enables the user to control and check on the thingies.

@@

With so many networking technologies available, care must be taken before picking one above another, however, we arrived at our choice, Wi-Fi, pretty fast. One of our primary requirements was that the network should exist in most homes already. This eliminates quite a few choices right away, leaving us with power lines, cellular networks such as 3G, phone line wiring, Bluetooth, and Wi-Fi. Power lines, which have the advantage of reaches all rooms in a home, have problems with the an uncertain network boundaries, since there is usually no special gateway or firewall preventing data from traveling to the neighbours power lines, unless extra hardware is added. Power lines are also inherently a wired network, which means an extra bridge device is needed between the power lines and wireless mobile devices if those are to interact with devices connected to it. The phone lines inside the home have similar problems, and are also far less ubiquitous in the home, with most homes only having a few phone plugs. Then there is Bluetooth, which most laptops and phones, even feature phones, come equipped with. Bluetooth is however limited in range. This leaves cellular networks, which are indeed wireless, but also owned and maintained by carrier companies, that charge fees for the data that travels on their network and not all areas are covered equally well. This makes cellular networks a less ideal choice, landing us squarely on Wi-Fi as the preferred option.

@@

The biggest drawback related to Wi-Fi is that it is power hungry compared to other technologies such as ZigBee \cite{zigbeeWebsite}. However, there is a general focus from the Wi-Fi radio chip makers to lower power consumption due to the massive surge of battery driven mobile devices the last few years, so the drawback will matter less as Wi-Fi radios become more power friendly. In addition, for many of the scenarios we have found users to be interested in, thingies will have a permanent power source and in the case where low powered sensor thingies are required, we expect these to use a protocol like ZigBee \cite{zigbeeWebsite} and be bundled with a special purpose thingy which act as a bridge between the individual ZigBee enabled thingies the Wi-Fi network. The bridge thingy could also perform other tasks so the extra cost would be negligible.

We also discussed the problems with oversaturation and general instabilities of the average home Wi-Fi network. This is of course very dependent on the model of the wireless access point used. Better models with good Wi-Fi radios and antennas, faster CPUs, and enough memory, can handle around 25 concurrent connected clients, while lesser models choke at at half as many connections. However, since one of our design goals was that it should be as inexpensive as possible to get started with smart home technology, we do not expect a average home Wi-Fi to get saturated right away. The scenario where a homeowner invests in over 20 new thingies at the same time might not be unlikely, but in that case we feel that investment in an extra access point to handle traffic from thingies is probably also within the budget. Some Wi-Fi experts claim that Wi-Fi clients also have a tendency to drop their connection from access points more often if using the higher bandwidth 802.11n and 802.11g protocols compared to the slower 802.11b. The solution here is to make thingies use 802.11b where possible, and since only a very small subset of thingies will require a 100\% stable connection, homeowners should experience no problems since thingies can simply reconnect if their connection drops.
\subsubsection{Communication protocols}

For communication, we use standard HTTP \cite{Fielding:1999:HTP:RFC2616}. This opens up for a huge list of possible controllers, basically any type of computing device or library that supports HTTP. This is especially important to us since we do not want to be limited to a certain framework, language, or platform, e.g. Java\footnote{\url{http://java.com}} or Android\footnote{\url{http://android.com}}. Other great properties of HTTP is that it is fairly lightweight and stateless protocol \cite{Fielding:1999:HTP:RFC2616}, which simplifies things considerably on the thingies. HTTP also supports our goal of an open infrastructure; it might very well be the best supported protocol among developers, and this enables the average app developer to build new functionality on top of the infrastructure on their platform of choice. Another benefit is that HTTP communication is very unlikely to be blocked by firewalls compared to other less used protocols, so there is a much bigger chance that things will just work.

However, HTTP's statelessness is also a weakness in common smart home scenarios, where a context change in a thingy should trigger an event on a controller. With basic HTTP, a controller waiting for a state change has to continuously poll the thingy it is monitoring to discover the state change in a timely manner. This is an waste of resources and could potentially deplete the resources of a thingy entirely, especially if more controllers are waiting for changes. Clearly, relying only on HTTP alone does not scale. Several solutions to this problem were considered, the biggest contender to be considered being multicast. The advantages of using multicast is that thingies can simply announce state changes on the network and it would not matter how many controllers were interested, the amount of resources required to do so would be the same. Ignoring the problems associated multicast on Wi-Fi, i.e. lost multicast packets are not retransmitted at the MAC layer \cite{conf/icnp/ChandraKMNPRR09}, the reason we did not choose multicast was that the protocol works on a lower level compared to HTTP, and that would limit the number of supported development platforms. We wanted to stay in the Web-stack of technologies so we chose the WebSocket protocol. In the recent years, the WebSocket protocol has been standardized by the Internet Engineering Task Force (IETF) \cite{IETFWebsockets} and is in the process of being standardized by the World Wide Web Consortium (W3C) \cite{w3cWebsockets}. The WebSocket protocol supports multiplexing bi-directional, full-duplex communications channels over a single TCP connection and is already supported by the latests versions of most major browsers\footnote{The WebSocket protocol is supported from Chrome 16, Firefox 11 and Internet Explorer 10.}, with numerous frameworks and libraries already built and more on the way. The WebSocket protocol was essentially designed to be a solution for the scenario where a server needs to push information to a client\cite{IETFWebsockets} which web developers have been struggling with for many years, so while it might not scale as well as multicast, we feel this is the best overall approach.

\subsubsection{Service Discovery}
Most home networks do not have traditional unicast DNS server set up, so we need a different way to discover thingies connect to the network, both during installation and also if and when their IP address changes. We have decided to use mDNS/Bonjour\footnote{Bonjour is available on \url{https://developer.apple.com/opensource/}} since it offers wide support across platforms and is relatively lightweight and open source.
\subsubsection{Security considerations}

The basic trust model for Thingies for Dummies is based on the security of the home's Wi-Fi. Assuming the Wi-Fi is probably configured using WPA2 with AES encryption and a reasonable passphrase, there is a clear boundary for the smart home, and an adversary would not be able to pick up data sent to and from thingies without being connected to the network. This also means, that anyone on the network has full control of the thingies attached to it, which can be a problem in some cases. We have discussed the possibility of adding an extra layer of security that would handle such a case. If the concern is only that unauthorised users cannot perform certain actions on a thingy, i.e. only change state, not update configuration, then we need a way to authenticate each user. This could be done using certificates that clients use to authenticate with. The certificate would be generated during the initial configuration of a thingy, and depending on the number of authorisation levels needed for a thingy, e.g. read only, update state, and update configuration, one or more unique certificates would be generated. Then it would be up to the controller devices used during the initial configuration to distribute the generated certificates to authorized users. However, it is not enough only to authenticate users, since the communication between users and thingies are still being transmitted over an open channel which can allow an adversary to perform replay attacks. The solution to this is to secure the communication channel between users and thinges, i.e. using HTTPS. Adding the extra level of security does raise the hardware requirements on thingies which increases cost and power consumption, so an environment that supports both ``open'', low risk thingies and ``secure'' thingies should provide the best of both worlds.

\subsection{Thingies - actuators and sensors}

@@

The first two points are important because thingies must be installable and manageable by less tech-savvy users. It is also reasonable to assume that a simple thingy with one or two functions is going to be more reliable than a device with several functions. Not allowing thingies to be dependent on other thingies or services simplifies everything, which again helps makes things easier for the less tech-savvy user. Making things simple also means being less dependent on constant firmware updates, which will reduce the amount of maintenance required after the initial installation. This is highly important, as Kindberg et al writes, \textit{ ``You can't reboot the world, let alone rewrite it, to introduce new functionality. Users shouldn't be led into a disposable physical world just because they can't keep up with software upgrades.''} \cite{kindbergSystem}.

Thingies can have running logic inside them, depending on their functionality, e.g. a power plug could be configured to turn on and off at specific times during the day, or turn on a secondary output if a specified amount of power is being drawn from a primary output, but it cannot depend on context outside of its domain, i.e. other thingies. Such logic should exist on controllers. Looking back at figure \ref{fig:infrastructureOverview}, recall that the controller is the users's smartphone.

@@

\begin{itemize}
\item Have a unique identifier
\item Have a passive NFC tag and/or a QR code with credentials for its bootstrap access point
\item Have a physical reset/install mode button
\item Be able to connect to an IP based network through a Wi-Fi radio
\end{itemize}

@@

The requirement for a Wi-Fi radio can be weakened to enable the scenario where a ZigBee enabled sensor thingies connect through a specialized ZigBee to Wi-Fi bridge thingy. We do however still require that all thingies can be reached on their own unique URI through the network. This is trivial for thingies connecting directly through Wi-Fi to the home network -- they simply have a standard URI \cite{uri-rfc} based on their IP address, e.g. \texttt{http://192.168.0.42}. In the bridge scenario, the bridged thingies, e.g. ZigBee sensors, may have to share the IP of the bridge. Luckily, the URI schema allows for ports to be defined, so each bridged thingy will be identified by the bridges IP and their own port, e.g. \lstinline{http://192.168.0.42:1337},\\\lstinline{http://192.168.0.42:1338}, etc.

@@

The structured description associates each state variable in the thingy with ID that is unique within each thingy. This makes it possible update the UI by filling in new state values without reloading the UI description. These IDs are also used when setting the state of the thingy. The following list summarizes the web service interface:

@@

The general bootstrapping process is intended to very simple from a user's point of view. First, the user plugs in and turns on the thingy. Then, the user starts the Home Remote app on his smartphone. When the app first runs, it will prompt the user for credentials to the home Wi-Fi network. We considered picking the current network automatically, but we want to ensure that thingies are connected to the right network, and the passphrase is not programmatically retrievable on Android\footnote{\url{http://developer.android.com/reference/android/net/wifi/WifiConfiguration.html}}. Then, the user selects ``Install new thingy''. The onscreen instructions will prompt the user to scan the QR code or hold the phone close to the thingy (if the phone is NFC enabled). From this point, the phone and the thingy are connected to each other, and the user is asked to give the new thingy a name so it is identifiable by humans. The phone also sends the Wi-Fi credentials to the thingy. The diagram in figure \ref{fig:installworkflow} illustrates the workflow of the bootstrapping protocol from a technical point of view.

@@

You cannot have a smart home without controllers, and to us, the obvious first choice for controllers are smartphones. It is a device which many users carry all the time and are comfortable using. As the previous sections show, we are by no means limited to smartphones -- any computing device supporting the HTTP protocol can act as a controller. This is also our only requirements for simple controllers, although we do require a full controller to be able to perform bootstrapping of new thingies, which is where smartphones really shine.

\subsubsection{End-user programming: Presets}
To enable the simplest form of user programming and automation on the controller level, and to make the scenario described in the introduction possible, we need a way to execute a batch of commands to different thingies. Our solution to this is something we have dubbed ``Presets''. A Preset consists of a name and a set of thingies, along with a desired state for each thingy. When the user defines a new Preset, he gives it a name and then adds one thingy at a time to the preset, specifying which state it should be in.When a Preset is triggered by the user, all thingies in the preset are put in the required state in sequential order, and the expected state of each thingies is checked. If any thingies in the Preset did not end up in the expected state, the user is notified. This also applies to thingies that are only sensors, where the Preset is simply used to verify their states. This enables relatively advanced scenarios such as our ``Lock Down'' scenario described earlier.

@@

The Internet of Things, implemented in e.g. HP's Cooltown project \cite{cooltown} is a concept related to ours in that it connects the physical and virtual worlds by creating an online presence for physical objects. The Cooltown project also mentions the use of local wireless networks as a means of access restriction.

In the following sections, we will relate our work to other smart home infrastructures as well as other technologies that move user interfaces between devices.

\subsection{Smart home technologies}
Our investigation of smart home technologies has not been as deep as we would have liked -- in part due to time constraints, but also because we did not have access to the relevant equipment. Nevertheless, a comparison follows.

X10\footnote{\url{http://x10.com}, \url{http://www.smarthome.com/about_x10.html}, and \url{http://kbase.x10.com/wiki/MyHouse_Online_Remote_Connection_Setup}} is a low-level protocol from the 70s running through the power lines in a home. It allows one to control a number of devices using a pre-defined set of commands\footnote{url{http://software.x10.com/pub/manuals/xtdcode.pdf}}. Thus, it requires a controller device which is connected to the power lines. When it comes to simplicity and end-user installation, X10 is probably the approach most similar to ours. However, there is the caveat that X10 signals can enter and leave a home through the power lines, so for a secure installation, an electrician is required to install a filtering device. Additionally, X10 does not have the convenience of wireless control built in; this requires the addition of a controller device that supports this.
